Conflict and Strategy in Collective Computation

### 2015-05-06 ms_draft_10_10_14.tex
writing ms and supplemental and making pretty figures

\documentclass{article}
\usepackage{latexsym}
\usepackage{amssymb,amsmath}
\usepackage{custom2}
\usepackage{graphicx} % for figures
\usepackage{epstopdf} % so can use EPS or PDF figures
\usepackage{caption}
\usepackage{subcaption}
\usepackage{url}
\usepackage{amssymb,amsfonts}
\usepackage[all,arc]{xy}
\usepackage{enumerate}
\usepackage{mathrsfs}
\usepackage{booktabs}
\usepackage[pdftex]{hyperref}
\usepackage{lscape}
\usepackage{xcolor}
\usepackage{natbib}

\captionsetup{justification=RaggedRight, singlelinecheck=false}
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
\newcommand{\argmax}{\text{argmax}}
\newcommand{\Tr}{\text{Tr}}
%\newtheorem{claim}{Claim}
\newtheorem{ass}{Assumption}

@@

\begin{document}
\begin{center}
\Large

\end{center}

\vspace{0pt}

\begin{center}
{\bf \LARGE{Conflict and Strategy in Collective Computation}}
\end{center}

\section{Abstract}
Groups of individuals making decisions in noisy environments perform collective computations in many biological systems, ranging in size from a brain made of billions of neurons to social groups made up of tens of monkeys. In each of these systems, pairs of individuals make decisions about their relative values. We use a leaky integrator model, a set of stochastic differential equations describing how decision variables change over time, to model the decision process. The decisions made between each pair form a directed network over the whole group. The degree of consensus in the group about each individual's value, as determined by the decision network, is a collective computation that is informative about the individuals' true values. We find that differences in the model when it is applied to neural and social systems do not affect the accuracy of decisions.  We find that conflicts of interest in the decision process affect individuals with low values the most and that conflicts of interest can incentivize these individuals to make more accurate decisions. We find that when pairwise decisions are accurate, local properties of the decision network are the most informative, but when accuracy decreases global properties are more informative. Finally, we find that the shape of the distribution of consensus scores is determined by the waiting costs of those decisions.

\section{Keywords }
collective computation, noisy learning, leaky integrator, decisions, social network, consensus, power, social structure

\section{Introduction} 
Many biological systems perform collective computation.  When presented with a set of moving dots, the brain, made up of billions of neurons, can decide whether the dots are moving left or right, even if the difference in the sizes of the groups moving either direction is not very large. A group of quorum sensing bacteria can sense its cell density and start to produce toxic, phosphorescent, or colorful molecules only when the density is sufficiently high. Flocks of birds and schools of fish can stay together and navigate with beautiful coordinated motion. In social groups of primates, the distribution of the degree of consensus about each monkeyÕs ability to use force is informative about the animalsÕ abilities and the animals use estimates of this emergent property to decide how to behave. In each of these examples, the individual members of the group make decisions based on noisy environmental information and a function of or algorithm acting on those individual decisions determines the groupÕs output. This two-step process allows the group to produce a successful output or behavior, despite computational limitations of its members and the conflicts of interest between them. While these examples come from quite different systems, the similarity in the phenomena suggest there may be common principles underlying collective computation in all of them.

 The sequential probability ratio test (SPRT) is the optimal way to decide between two alternatives in a noisy environment in the sense that for a given amount of time the SPRT is the most accurate algorithm and for a given error rate the SPRT reaches a decision most quickly.  The algorithm works by keeping track of the amount of evidence supporting one alternative and deciding in favor of that alternative once that amount is sufficiently high or deciding against it once that amount is sufficiently low. While this algorithm assumes the decider has perfect memory, in reality this is rarely the case and a simple way to model memory loss is to assume that the amount of evidence decays back toward 0 at teach point in time. This forgetful model of decision-making can be transformed into a leaky integrator model, a set of stochastic differential equations (SDEs) describing how the variables accumulating evidence for each alternative change over time. The random walk described by the SDEs is biased by the true state of the environment. This model is applicable to any system in which an individual with imperfect memory gathers evidence and chooses the alternative for which he has gathered more support.  The leaky integrator model, while a powerful for modeling noisy learning dynamics, has been applied rather phenomenologically, and we seek here to justify its use with a mechanistic derivation.

In this paper, we will study a group of individuals, each pair of which is trying to make a decision about their relative values. Each individual will have a decision variable indicating how much evidence it has accumulated and a threshold.  If the decision variable reaches the threshold, the individual will signal that a decision has been made. Both individuals involved in the decision value making a correct and fast decision. Since it takes time to gather enough evidence to make a correct decision, there is a tradeoff between accuracy and speed. Additionally, each individual might prefer that the decision be made in their favor, whether or not they truly have a higher value.  This introduces a conflict of interest. The thresholds are the strategies they use to negotiate accuracy, speed, and their personal preferences.

The leaky integrator model has been applied to neural decision-making, and it successfully explains and predict performance in a number of experiments  \citep{Eckhoff:2008uq, Brown:2005fk,Feng:2009kl,Bogacz:2006uq}. In this application, each of two neural populations responds to evidence for either left or right motion.  The SDEs describe how the firing rates of the two populations change over time and a decision is made once one the difference in firing rates is sufficiently high. Usually, the visual stimulus driving this decision process is a set of dots, some percentage of which are moving coherently either left or right and the rest of which are moving randomly.  The bias in the SDEs is in favor of the direction in which the coherent dots are moving.  However, this process could be used by neural populations deciding between any two possibilities, for example between up and down motion or between two colors, and a group of neural populations could be used to distinguish between or rank a whole set of possibilities.

While the SPRT has been applied to animal conflicts  \citep{Froment:2010fk}, to our knowledge the leaky integrator model has not been. In this application, each of two animals has a belief about its dominance with respect to the other which increases or decreases as the animal wins or loses fights against its opponent. The SDEs describe how these beliefs change over time. The source of evidence is the fight history between the two animals. Even if one animal is stronger and tends to win fights, the outcome of a fight is affected by the identity and presence of each animal's allies, motivation levels,  external environmental variables, and other stochastic factors, which introduce noise.   The bias in the SDEs is in favor of the animal that is more likely to win. Once an animals' opinion goes below a threshold, it decides to agree to be subordinate to the other. A correct decision is one in which the weaker individual makes this agreement.  In pigtailed macaques ({\it Macaca nemestrina}) and other species of macaques, there is a formal subordination signal that an animal can use to communicate its agreement to the subordinate role in their relationship \citep{Flack:2007kx, Flack:2006fk,Flack:2004oq, Waal:1985fk,Caldecott:1986uk,Preuschoft:2004ly}.  This signal marks the end of the decision process.

The leaky integrator model can thus be used to describe a decision being made between each pair of individuals in a group.  These decisions form a directed network, each edge of which reflects a decision about the relative values of the nodes it connects. In many systems, the degree of consensus in a group about each individuals' state is informative about the individuals'  true states and affects the individuals' functional behavior \cite{Brush:2013fk}. In previous work, we developed a number of network metrics to measure the degree of consensus in the group about each node and we found a few that produced scores that were predictive of individuals' functional behaviors \cite{Brush:2013fk,Flack:2006uq}. The output of one of these consensus measures being applied to a decision network would be a set of consensus scores, a collective computation which would be informative about the individuals' true states.

In the social case, this consensus is referred to as social power. In macaques, power is a useful variable since it reflects the animals' true abilities and the animals themselves do use estimates of their peers' and their own power to decide how to behave \cite{Flack:2004oq,Flack:2006fk,Flack:2006uq}.  The distribution of power scores in an empirical group of pigtailed macaques is heavy tailed and this allows the individuals in the tail to perform different functions in the social group than the rest of its members \cite{Flack:2005uq,Flack:2006fk,Flack:2006uq}.  Further, the shape of the power distribution varies across species of macaques \cite{Caldecott:1986uk,Flack:2004oq,Preuschoft:2004ly,Waal:1985fk,Thierry:2000ij}, and it is an open question whether this variation can be explained by differences in the underlying distributions of fighting abilities, by differences in the way animals in different species resolve their differences, or through other mechanisms. This kind of consensus may be important in neural systems, as well.  A number of neural populations may respond to or encode different environmental stimuli or variables.  The consensus in a group of neural populations about the value of a number of environmental stimuli could help the brain make decisions about the relative values of many possibilities.

There are thus three factors affecting the collective computation: the statistical properties of the environment, the strategies individuals use to make decisions about their relative values, and the method of measuring consensus from a network of decisions.  In this paper, first we derive the leaky integrator model from a simple mechanistic description of the accumulation of information, making two simplifying assumptions about the timescales on which probabilistic events occur \cite{Gillespie:2000fk}. We then use this model to answer the following questions. Do differences in the decision making model as it is applied to neural and social systems affect the accuracy of the decision process? What threshold strategies should individuals with conflicting interests use to make decisions in a group context? How do the individualsÕ threshold strategies affect the collective computation? What function of the decision network produces the most informative collective computation? How well does our model describe empirical social systems?

\section{Model} 
\subsection{Stochastic differential equations }    
%% make debt to Gillespie clearer

There are two possible types of environmental evidence: in the neural system, a dot can move either left or right and in the social system, a fight can be won by one animal or another.  For clarity in the following we will refer to these left or right dots, but the biological interpretation of the two types of inputs depends on the system. There are two decision variables, $X_1$ and $X_2$. In the absence of any input, the decision variables leak back towards $0$ with rate $\ell$.  (A table of all variables used in the text is given in Table \ref{variables}.)  If there is no input, then over a period length $\tau$ a decision variable decreases as $X(t+\tau)=(1-\ell\tau)X(t)$.

Each decision variable responds positively to one type of input and negatively to the other.  In the neural case, one neural population responds positively to left dots and negatively to right dots and the other neural population does the opposite.  In the social case, one animal's estimate of its own dominance increases when it wins fights and decreases when it loses fights, and the other animal's estimate does the opposite. Neither decision variable has perfect memory.  Specifically, $X_1$ increases by a boost $b$ when it sees a left dot and decrements by $-b$ when it sees a right dot, and conversely for $X_2$.     
To calculate the variables at time $t+\tau$, we  count how many times each of type of input occurred in the time since $t$ and add the changes resulting from these events to the background leaky estimate (as in \cite{Gillespie:2000fk}):
\begin{align*}
X_1(t+\tau)&=(1-\ell\tau)X_1(t)+b\times\# \text{ of left dots in }[t,t+\tau)-b\times\# \text{ of right dots in }[t,t+\tau)
\\ X_2(t+\tau)&=(1-\ell\tau)X_2(t)-b\times\# \text{ of left dots in }[t,t+\tau)+b\times\# \text{ of right dots in }[t,t+\tau). 
\end{align*}

Our first assumption is that the rates at which the two types of inputs occur is constant over time.  We can thus describe the number of each type of event with a Poisson random variable, $N_\text{L}$ and $N_\text{R}$.  We rewrite our basic equations to give 
\begin{align*}
X_1(t+\tau)&=(1-\ell\tau)X_1(t)+bN_\text{L}-bN_\text{R}
\\ X_1(t+\tau)&=(1-\ell\tau)X_2(t)-bN_\text{L}+bN_\text{R}.
\end{align*}
If events happen at a rate $r$ and the event is a left dot with probability $c$ and a right dot with probability $1-c$, then the expectation of $N_L$ and $N_R$ in a period of length $\tau$ are, respectively, $\tau r c$ and $\tau r(1-c)$. In the neural case, $c$ is related to the ``coherence'' of the visual stimulus.  A $c$ close to $0$ or $1$ means that most of the events are of one type or the other and the decision is easier to make than when $c$ is close to $.5$.

If enough events happen in the period of time from $t$ to $t+\tau$ then we can approximate the Poisson random variables with Normal random variables with mean and variance equal to the mean of the Poisson random variables.  Our second assumption, then, is that the period of time of length $\tau$ is long enough to make this approximation. Let $Z_\text{L}$ and $Z_\text{R}$, be independent standard Normal random variables, i.e. with mean $0$ and standard deviation $1$.  Then our equations for $X_i(t+\tau)$ become
\begin{align*}
X_1(t+\tau)&=(1-\ell\tau)X_1(t)+b\bigg(\tau rc+\sqrt{\tau rc}Z_{\text{L}}\bigg)-b\bigg(\tau r(1-c)+\sqrt{\tau r(1-c)}Z_{\text{R}}\bigg)
\\X_2(t+\tau)&=(1-\ell\tau)X_2(t)-b\bigg(\tau rc+\sqrt{\tau rc}Z_{\text{L}}\bigg)+b\bigg(\tau r(1-c)+\sqrt{\tau r(1-c)}Z_{\text{R}}\bigg).
\end{align*}
Finally, as we make the period of time shorter and shorter, making $\tau$ infinitesimally small, these equations become stochastic differential equations,
\begin{equation}
\begin{array}{ll}
dX_1&=\bigg(-\ell X_1(t)+br(2c-1)\bigg)dt+\bigg(b\sqrt{rc}\bigg)dW_\text{L}t-\bigg(b\sqrt{r(1-c)}\bigg)dW_\text{R}t
\\dX_2&=\bigg(-\ell X_2(t)-br(2c-1)\bigg)dt-\bigg(b\sqrt{rc}\bigg)dW_\text{L}t+\bigg(b\sqrt{r(1-c)}\bigg)dW_\text{R}t
\end{array}
\end{equation}
where $dW_{\text{L}}$ and $dW_{\text{R}}$ are independent Brownian motions representing, respectively, the left and right inputs.  The assumptions about the timescales on which inputs occur are particularly reasonable in the social system, although the successful application of this type of model to neural populations implies they are not unreasonable in that system.  In Table \ref{variables}, we list the inputs, outputs, and variables of the decision model and how they should be interpreted in the neural and social systems.

\subsection{Reaching a decision }
Each decision variable, $X_1$ and $X_2$, accumulates information about the stimulus.  The sign and size of the two variables indicate how much evidence there is for each possible decision outcome.  When a lot of evidence for the first possibility has been observed, $X_1$ will be large and positive, and similarly for the second possibility.  If the difference in the variables can be observed, the difference indicates the relative strengths of the evidence for either possibility. If $Y=X_1-X_2$, when $Y$ is large enough the system decides on $X_1$ and when $Y$ is low enough the system decides on $X_2$.  Specifically, there are two thresholds, $T_2$ and $T_1$ such that if $Y>T_1$ the decision is for $X_1$ and if $Y<-T_2$ the decision is for $X_2$.  However, it may be the case that the difference in the variables cannot be observed.  In this case, the decision depends on whether $X_1$ or $X_2$ hits a threshold first.  Again, there are two thresholds, $T_2$ and $T_1$, and if $X_1<-T_1$ the decision is for $X_1$ and if the $X_2<-T_2$ the decision is for $X_2$.

In the neural system, the two stochastic equations describing the activity of the two neural populations integrating environmental evidence are often reduced to the one-dimensional equation describing the difference in the activity levels to make the system easier to analyze \cite{Brown:2005fk,Bogacz:2006uq,Feng:2009kl}.  However, in the social system we are considering, we cannot make any assumptions about the existence of a third party evaluating the difference in the opinions of the two animals.  A ``decision," i.e. the emission of a signal from one of the two animals, is only reached when one of the two animals' opinions goes below a certain threshold.  The social system is therefore inherently two-dimensional, whereas the neural system may be adequately described by one dimension.  In Table \ref{differences}, we compare how the decision is made and how the decision process is optimized in the two systems.

\subsection{Utility of the decision }
The thresholds affect how long the decision takes and the probabilities with which each output is decided upon.  A good decision is one that reaches the correct output quickly, i.e. the expected time until a decision is reached (decision time, DT) and the probability that an incorrect decision is made (error rate, ER) are low.   However, it is impossible to minimize both simultaneously since waiting longer and accumulating more evidence makes the decision more accurate and conversely.  It may also be the case that each individual prefers one output over the other, regardless of which is ``correct''.  In the social case, an animal would prefer that its partner agreed to be subordinate, whether or not it is truly stronger.  In the neural case, different neural populations may receive different rewards if the decision reaches different outputs, regardless of which of the alternatives is ``true."  In these cases, a variable wants to maximize the probability that his preference is reached (probability of preference being chosen, $\text{PP}_i$).

To describe the tradeoffs between error rate, decision time, and preference, we quantify the utility of the decision process by introducing three weights, $w_1$, $w_2$, and $w_3$ such that $w_1+w_2+w_3=1$.  These weights describe how the three quantities are prioritized.  For individual $i$, we define
\begin{equation*}
U_{i}=w_1\text{ER}+w_2\text{DT}+w_3(1-\text{PP}_i).
\end{equation*}
The decision is better if $U_i$ is lower.  In the social system, $w_2$ can be interpreted as the cost of fighting since when $w_2$ is higher, the time spent fighting until a decision is reached is more costly.  Similarly, $w_3$ can be interpreted as the benefit from being the dominant animal in a pair. In both cases, $w_3$ represents how stubborn the individuals are about their preferences. If $w_3\gneq 0$ and the individuals prefer different outcomes, neither can simultaneously be satisfied since $\text{PP}_1=1-\text{PP}_2$.

As show in the Appendix, each of the three decision properties, ER, DT, and PP satisfies a partial differential equation that depends on the SDEs and the parameters of the model. There are analytical solutions to these equations when the system of SDEs is reduced to one dimension. However, we were not able to find an analytical solution for the full two-dimensional system and we therefore used numerical methods to solve the PDEs in this case.

\subsection{Optimal thresholds }
An individual-- either a neural population or a monkey-- can be engaged in either a single decision or in many decisions at once.  In the social system, for example, an animal wants to decide whether it is dominant or subordinate to every other member of its social group.  For any pair of individuals, each has a utility, $U_1$ and $U_2$, and consequently each has a Nash equilibrium threshold such that, if they choose those thresholds, neither has an incentive to choose another.  Similarly, in a group of individuals, each individual $i$ has a utility $U_{ij}$ from its decision process with individual $j$.  His total utility is given by the average of these, $\langle U_{ij}\rangle _j$. These lead to an optimized set of thresholds such that no individual has an incentive to choose another threshold.

>>>>

\documentclass{article}
\usepackage{latexsym}
\usepackage{amssymb,amsmath,amsfonts}
\usepackage{custom2}
\usepackage{graphicx} 
\usepackage{epstopdf} 
\usepackage{caption}
\usepackage{subcaption}
\usepackage{url}
\usepackage[all,arc]{xy}
\usepackage{enumerate}
\usepackage{mathrsfs}
\usepackage{booktabs}
\usepackage[pdftex]{hyperref}
\usepackage{lscape}
\usepackage{xcolor}
\usepackage{natbib}

\captionsetup{justification=RaggedRight, singlelinecheck=false}
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
\newcommand{\argmax}{\text{argmax}}
\newcommand{\Tr}{\text{Tr}}
\newtheorem{ass}{Assumption}

@@

\title{Conflict and Strategy in Collective Computation}
\author{Eleanor Brush, David Krakauer, Jessica Flack}

\begin{document}
\maketitle

\section*{Abstract}
Collective computations occur in many biological systems, ranging in size from a brain made of billions of neurons to a social group made up of tens of monkeys. In each of these systems, a group-level output emerges from the behavior of individuals making decisions in noisy environments. In particular, in the brain and in primate social groups, pairs of individuals must make decisions about their relative relevance or ability. We use a leaky integrator model, a set of stochastic differential equations describing the dynamics of two decision variables, to model this decision process. The decisions made between each pair form a directed network over the whole group. The degree of consensus in the group about each individual's value, as determined by the decision network, is a collective computation that is informative about the individuals' true values. First, we show that the same model can be used to describe both neural and social systems. Next, we find that conflicts of interest in the decision process most strongly affect individuals at the bottom of the group and that conflicts of interest can incentivize these individuals to make more accurate decisions. Then, we find that when pairwise decisions are accurate, local properties of the decision network are the most informative, but global properties are more informative when decisions are less accurate or when the decision network is not fully formed. Finally, we find that the shape of the distribution of consensus scores is determined by the waiting costs of those decisions.

\section*{Keywords }
collective computation, noisy learning, leaky integrator, decisions, social network, consensus, power

\section*{Introduction} 
Many biological systems perform collective computation.  A group of quorum sensing bacteria can sense its density and start to produce toxic, phosphorescent, or colorful molecules once the density is sufficiently high. Flocks of birds and schools of fish can stay together and navigate with beautiful coordinated motion. When presented with a set of moving dots, the brain, made up of billions of neurons, can decide whether more dots are moving left or right, even when the numerical superiority of one direction is very small. In social groups of primates, the distribution of the degree of consensus about each monkey's ability to use force is informative about the animalsÕ abilities and the animals use estimates of this emergent property to decide how to behave. In each of these examples, the individual members of the group make decisions based on noisy environmental information and the collective computation emerges from those individual decisions. Through this two-step process the group produces a successful output or behavior, despite the computational limitations of its members and the conflicts of interest between them. While these examples come from quite different systems, the similarity in the phenomena suggest there may be common principles underlying collective computation in all of them. (In Table \ref{examples} we identify the components of the collective computation--- the environmental information, the individual, the decision that each individual must make, and the resulting collective computation---in each of these systems.)

%Both individuals involved in the decision value making a correct and fast decision. Since it takes time to gather enough evidence to make a correct decision, there is a tradeoff between accuracy and speed. Additionally, each individual might prefer that the decision be made in their favor, whether or not they truly have a higher ability.  This introduces a conflict of interest. Each individual will have a strategy to negotiate accuracy, speed, and their personal preferences.

To understand these collective computations, we must understand how the individuals make decisions. The sequential probability ratio test (SPRT) is the optimal way to decide between two alternatives in a noisy environment, in the sense that, for a given amount of time, the SPRT is the most accurate algorithm and, for a given error rate, the SPRT reaches a decision most quickly.  The algorithm works by keeping track of the amount of evidence supporting each alternative and deciding in favor of one once the evidence for it is sufficiently high. This algorithm assumes the individuals making the decision have perfect memory, but in reality this is rarely the case.  A simple way to model memory loss is to assume that the amount of evidence decays back toward zero at each point in time. This forgetful model of decision-making can be transformed into a leaky integrator model, a set of stochastic differential equations (SDEs) describing how the variables accumulating evidence for each alternative change over time. The random walk described by the SDEs is biased by the true state of the environment. This model is applicable to any system in which an individual with imperfect memory gathers evidence and chooses the alternative for which it has gathered more support.

The leaky integrator model has been applied to neural decision-making, and it successfully predicts performance in a number of experiments  \citep{Eckhoff:2008uq, Brown:2005fk,Feng:2009kl,Bogacz:2006uq}. In this system, each of two neural populations responds to evidence for either left or right motion.  The SDEs describe how the firing rates of the two populations change over time and a decision is made once one the difference in firing rates is sufficiently high. The visual stimulus driving this decision process is usually a set of dots, some percentage of which are moving coherently either left or right and the rest of which are moving randomly.  The bias in the SDEs is in favor of the direction in which the coherent dots are moving.  The model could also describe neural populations deciding between other possibilities, for example between up and down motion or between two colors. In order to distinguish between or rank a whole set of possibilities, the brain could rely on a group of neural populations, each responding to one of the options.

While the SPRT has been applied to animal conflicts  \citep{Froment:2010fk}, to our knowledge the leaky integrator model has not been. In this application, each of two animals has an opinion about its dominance with respect to the other, which goes up or down as the animal wins or loses fights against its opponent. Even if one animal is stronger, the outcome of a fight is affected by the identity and presence of each animal's allies, their motivation levels,  external environmental variables, and other stochastic factors, which introduce noise that is captured by the SDEs.   The bias in the model is in favor of the animal that is more likely to win. Once an animal's opinion goes below a threshold, it decides to agree to be subordinate to the other. A correct decision is one in which the weaker individual makes this agreement.  In pigtailed macaques ({\it Macaca nemestrina}), there is a formal subordination signal that an animal can use to communicate its agreement to the subordinate role in their relationship \citep{Flack:2007kx, Flack:2006fk,Flack:2004oq, Waal:1985fk,Caldecott:1986uk,Preuschoft:2004ly}.  This signal marks the end of the decision process.

In both systems, the decisions between pairs of individuals form a directed network, each edge of which reflects a decision about the relative values of the nodes it connects. In many systems, the degree of consensus in a group about each individual's value is informative about the individuals'  true values and affects the individuals' functional behaviors \cite{Brush:2013fk}. In previous work, we developed a number of network metrics to measure the degree of consensus in a group about each individual and we identified a set of algorithms that produced scores that were predictive of individuals' functions in a group in multiple empirical systems \cite{Brush:2013fk,Flack:2006uq}. The consensus scores that result from applying one of these consensus measures to the decision network is the collective computation that we are interested in. There are three factors affecting this collective computation: the statistical properties of the environment, the strategies individuals use to make decisions about their relative values, and the method of measuring consensus from a network of decisions.

In the social system, this consensus is referred to as social power. In macaques, power is a useful variable since it reflects the animals' true fighting abilities and the animals themselves use estimates of their peers' and their own power to decide how to behave \cite{Flack:2004oq,Flack:2006fk,Flack:2006uq}.  The distribution of power scores in a group of pigtailed macaques is heavy tailed and this allows the individuals in the tail to perform different functions  than the rest of the animals \cite{Flack:2005uq,Flack:2006fk,Flack:2006uq}.  The shape of the power distribution varies across species of macaques \cite{Caldecott:1986uk,Flack:2004oq,Preuschoft:2004ly,Waal:1985fk,Thierry:2000ij}, and it is an open question whether this variation can be explained by differences in the underlying distributions of fighting abilities, by differences in the way animals in different species resolve their differences, or by other mechanisms. Consensus may be important in neural systems, as well.  A number of neural populations may respond to or encode different environmental stimuli or variables.  The consensus in a group of neural populations about the value of a number of environmental stimuli could help the brain make decisions about the relative values of many possibilities.

In this paper, we derive the leaky integrator model from a simple mechanistic description of the accumulation of information, making two simplifying assumptions about the timescales on which probabilistic events occur \cite{Gillespie:2000fk}. While the model has been successfully used to describe neural learning dynamics, its application has been rather phenomenological and this mechanistic derivation justifies its application in any system where the assumptions are reasonable. We then use the leaky integrator model to answer the following questions. Do differences in the decision making model as it is applied to neural and social systems affect the accuracy of the decision process? What strategies should individuals with conflicting interests use to make decisions in a group context? How do the individuals' strategies affect the collective computation? What function of the decision network produces the most informative collective computation? How well does our model describe empirical social systems?

\section*{Model} 
\subsection*{Stochastic differential equations }    
\ In chemical systems, stochastic differential equations are used to describe the dynamics of the concentrations of various solutes, and these Langevin equations can be derived from a mechanistic description of chemical reactions \cite{Gillespie:1992vn,Gillespie:2000fk}. However, the SDEs used to model noisy decision processes are usually applied without any mechanistic justification. By following the mathematical derivation of the SDEs in chemical systems given by Gillespie \cite{Gillespie:2000fk},  we can derive the SDEs used in the leaky integrator model.

In a decision between two alternatives, there are two decision variables, $X_1$ and $X_2$, each indicating the evidence accumulated for one of the options.  In the absence of any input, the decision variables leak back towards $0$ with rate $\ell$.  (A table of all variables used in the text is given in Table \ref{variables}.)  If there is no input, then over a period length $\tau$ a decision variable decreases as $X(t+\tau)=(1-\ell\tau)X(t)$. If there is input, each decision variable will incorporate the new evidence by responding positively to one type of input and negatively to the other.  Specifically, $X_1$ increases by an amount $b$ when it sees evidence for option $1$ and decrements by $-b$ when it sees evidence for option $2$, and conversely for $X_2$.  In the neural case, one neural population responds positively to left dots and negatively to right dots and the other neural population does the opposite.  In the social case, one animal's estimate of its own dominance increases when it wins fights and decreases when it loses fights, and the other animal's estimate does the opposite.  For clarity, in the following we will refer to the inputs as left or right dots, but they could also be referred to as wins or losses. To calculate the variables at time $t+\tau$, we  count how many times each of type of input occurred in the time since $t$ and add the changes resulting from these events to the background leaky estimate (as in \cite{Gillespie:2000fk}):
\begin{align*}
X_1(t+\tau)&=(1-\ell\tau)X_1(t)+b\times\# \text{ of left dots in }[t,t+\tau)-b\times\# \text{ of right dots in }[t,t+\tau)
\\ X_2(t+\tau)&=(1-\ell\tau)X_2(t)-b\times\# \text{ of left dots in }[t,t+\tau)+b\times\# \text{ of right dots in }[t,t+\tau). 
\end{align*}

Our first assumption is that the rates at which the two types of inputs occur is constant over time.  We can thus describe the number of each type of event with a Poisson random variable, $N_\text{L}$ and $N_\text{R}$, giving 
\begin{align*}
X_1(t+\tau)&=(1-\ell\tau)X_1(t)+bN_\text{L}-bN_\text{R}
\\ X_1(t+\tau)&=(1-\ell\tau)X_2(t)-bN_\text{L}+bN_\text{R}.
\end{align*}
If events happen at a rate $r$ and the event is a left dot with probability $c$ and a right dot with probability $1-c$, then the expectation of $N_L$ and $N_R$ in a period of length $\tau$ are, respectively, $\tau r c$ and $\tau r(1-c)$. In the neural case, $c$ is related to the ``coherence'' of the visual stimulus.  A value of $c$ close to $0$ or $1$ means that most of the events are of one type or the other and the decision is easier to make than when $c$ is close to $.5$.

If enough events happen in the period of time from $t$ to $t+\tau$ then we can approximate the Poisson random variables with normal random variables with mean and variance equal to the mean of the Poisson random variables.  Our second assumption, then, is that the period of time of length $\tau$ is long enough to make this approximation. Let $Z_\text{L}$ and $Z_\text{R}$, be independent standard Normal random variables, i.e. with mean $0$ and standard deviation $1$, giving
\begin{align*}
X_1(t+\tau)&=(1-\ell\tau)X_1(t)+b\bigg(\tau rc+\sqrt{\tau rc}Z_{\text{L}}\bigg)-b\bigg(\tau r(1-c)+\sqrt{\tau r(1-c)}Z_{\text{R}}\bigg)
\\X_2(t+\tau)&=(1-\ell\tau)X_2(t)-b\bigg(\tau rc+\sqrt{\tau rc}Z_{\text{L}}\bigg)+b\bigg(\tau r(1-c)+\sqrt{\tau r(1-c)}Z_{\text{R}}\bigg).
\end{align*}
Finally, as we make the period of time shorter and shorter, making $\tau$ infinitesimally small, these equations become stochastic differential equations,
\begin{equation}
\begin{array}{ll}
dX_1&=\bigg(-\ell X_1(t)+br(2c-1)\bigg)dt+\bigg(b\sqrt{rc}\bigg)dW_\text{L}t-\bigg(b\sqrt{r(1-c)}\bigg)dW_\text{R}t
\\dX_2&=\bigg(-\ell X_2(t)-br(2c-1)\bigg)dt-\bigg(b\sqrt{rc}\bigg)dW_\text{L}t+\bigg(b\sqrt{r(1-c)}\bigg)dW_\text{R}t
\end{array}
\end{equation}
where $dW_{\text{L}}$ and $dW_{\text{R}}$ are independent Brownian motions representing, respectively, the left and right inputs.  The assumptions about the timescales on which inputs occur are reasonable in the social system, and the successful application of this type of model to neural populations implies they are not unreasonable in that system.  In Table \ref{variables}, we list the inputs, outputs, and variables of the decision model and how they should be interpreted in the neural and social systems.

\subsection*{Reaching a decision }\ There are two ways to use the decision variables $X_1$ and $X_2$ to make a decision.  If the difference in the variables can be observed, it indicates the relative strengths of the evidence for each option. If $Y=X_1-X_2$, the system should decides on $X_1$ when $Y$ is large and positive and on $X_2$ when $Y$ is large and negative.  Specifically, there are two thresholds, $T_1$ and $T_2$ such that if $Y>T_1$ the decision is for $X_1$ and if $Y<-T_2$ the decision is for $X_2$.  However, it may be the case that the difference in the variables cannot be observed.  In this case, the decision depends on whether $X_1$ or $X_2$ hits a threshold first.  Again, there are two thresholds, $T_1$ and $T_2$, and if $X_2<-T_2$ the decision is for $X_1$ and if the $X_1<-T_1$ the decision is for $X_2$.

In the neural system, the two stochastic equations describing the activity of the two neural populations integrating environmental evidence are often reduced to the one-dimensional equation describing the difference in the activity levels \cite{Brown:2005fk,Bogacz:2006uq,Feng:2009kl}.  The reduced model is easier to analyze, but there is no empirical evidence for a mechanism in the brain that measures the difference in activity rates and the full two-dimensional model should apply equally well to neural decision making, a possibility we consider here. In the social system, we cannot make any assumptions about the existence of a third party evaluating the difference in the opinions of the two animals.  A decision, i.e. the emission of a signal from one of the two animals, is only reached when one of the two animals' opinions goes below its threshold.  The social system is therefore inherently two-dimensional.  In Table \ref{differences}, we compare how the decision is made and how the decision process is optimized in the two systems.

\subsection*{Utility of the decision }
\ The thresholds affect how long the decision takes and the probabilities with which each output is decided upon.  A good decision is one that reaches the correct output quickly, i.e.\ the expected time until a decision is reached (decision time, DT) and the probability that an incorrect decision is made (error rate, ER) are low.   However, it is impossible to minimize both simultaneously since waiting longer and accumulating more evidence makes the decision more accurate and conversely.  It may also be the case that each individual prefers one output over the other, regardless of which is correct.  In the social case, each animal would prefer that its partner agree to be subordinate, whether or not it is truly stronger.  In the neural case, different neural populations may receive different rewards if the decision reaches different outcomes, regardless of which of the alternatives is correct.  Therefore, an individual may want to maximize the probability that its preferred outcome is reached (probability of preference being chosen, $\text{PP}_i$).

To describe the tradeoffs between error rate, decision time, and preference, we quantify the utility of the decision process by introducing three weights, $w_1$, $w_2$, and $w_3$ such that $w_1+w_2+w_3=1$.  These weights describe how the three quantities are prioritized.  For individual $i$, we define
\begin{equation*}
U_{i}=w_1\text{ER}+w_2\text{DT}+w_3(1-\text{PP}_i).
\end{equation*}
The decision is better if $U_i$ is lower.  In the social system, $w_2$ can be interpreted as the cost of fighting since when $w_2$ is higher, the time spent fighting until a decision is reached is more costly.  In the neural system, $w_2$ depends on whatever costs the brain or the whole animal incurs by waiting for a decision. Given that $\text{PP}_1=1-\text{PP}_2$, it is impossible for them both to maximize $\text{PP}_i$, so the personal preference weight $w_3$ indicates the strength of the conflict of interests between individuals and how stubborn the individuals are about their preferences. In the social system, $w_3$ can be interpreted as the benefit from being the dominant animal in a pair.

As show in the Supplementary Information, each of the three decision properties, ER, DT, and PP satisfies a partial differential equation that depends on the SDEs and the parameters of the model. There are analytical solutions to these equations when the system of SDEs is reduced to one dimension. However, we were not able to find an analytical solution for the full two-dimensional system and we therefore used numerical methods to solve the PDEs in this case.

\subsection*{Nash equilibrium thresholds }
\ In a group of individuals, each with a value $a_i$, the difficulty with which the pair $i,j$ makes a decision increases with the difference in value, i.e. $c_{ij}$ is an increasing function of $|a_i-a_j|$. In the social system, the value $a_i$ reflects an animal's fighting ability. In the neural system, the value $a_i$ is high if the property (e.g.\ left movement or red color) a neural population responds to is abundant in the visual stimulus. We assume that each individual has the same decision threshold for all the decisions processes with each of its peers and, given those thresholds, each individual $i$ has a utility $U_{ij}$ from its decision process with individual $j$ and a total utility given by the average of these, $\langle U_{ij}\rangle _j$. For each set of values $\{a_1,\dots,a_N\}$, we find the Nash equilibrium thresholds such that no individual has an incentive to choose another threshold to improve its utility.  Since the Nash thresholds depend on the values $\{a_1,\dots,a_N\}$,  we repeatedly draw a set of values from a uniform distribution and find the Nash thresholds for each set. Then we find the average Nash threshold as a function of an animal's rank in the set of fighting abilities.
